A Sound Deep Embedding of Arbitrary Normal Modal Logics in Coq

### 2022-09-04 paper/paper.tex
Fix some typos

\begin{itemize}
  \item We define a \emph{deep embedding} of modal logic syntax in Coq and
    define its Kripke semantics.
  \item We model the Hilbert deductive systems as a Coq inductive type that is
    parameterized by different axiom systems like K, D, B, T, K4, K5, S4, S5 and GL,
    and prove its soundness with respect to Kripke semantics.
    We also formalize several frame correspondence results.
  \item We show a proof of L\"ob's theorem using the Hilbert deductive system
    developed in our library.
\end{itemize}

@@

Coq is a proof assistant based on the calculus of inductive
constructions (CIC)~\cite{manual_cic}, a higher-order typed
$\lambda$-calculus extended with inductive definitions.  Theorem
proving in Coq follows the ideas of the so-called
``BHK-cor\-res\-pon\-dence''\footnote{Abbreviation of Brouwer, Heyting,
  Kolmogorov, de Bruijn and Martin-L\"of Correspondence. This is also
  known as the Curry-Howard isomorphism.}, where types represent
logical formulas, $\lambda$-terms represent proofs
\cite{Bertot10} and the task of checking if a piece of text is a
proof of a given formula corresponds to checking if the term that
represents the proof has the type corresponding to the given formula.

However, writing a proof term whose type is that of a logical formula
can be a hard task, even for very simple propositions.  In order to
make the writing of complex proofs easier, Coq provides
\emph{tactics}, which are commands that can be used to construct proof
terms in a more user-friendly way.

@@

\paragraph{An overview of modal logic.} Modal logics are designed for reasoning
about truth across various --- abstract --- worlds. In such logics, a
proposition may be true in some world, but false in another one. The versions of
modal logics considered in this work extends traditional propositional logics
with two operators on propositions: $\Box$ (box) and $\Diamond$ (diamond).
The syntax of modal logic formulas are defined by the following context free
grammar (where $\circ \in\{\land,\lor,\to,\leftrightarrow\}$ and
$\mathcal{P} = \{A,B,C,...\}$ is the enumerable set of propositional variables):
\[
  \begin{array}{lcl}
    \varphi & ::= & A\,|\,\neg\varphi\,|\,\varphi\circ\varphi\,|\,\Box\varphi\,|\,\Diamond\varphi
  \end{array}  
\]
Intuitively, the semantics of the operator $\Box$ is similar to an universal
quantifier, since it express \emph{necessity}: $\Box\,\varphi$ is true at the
current world iff $\varphi$ is true at all worlds accessible from the current
world. Similarly, the semantics of operator
$\Diamond$ is that a formula $\Diamond\varphi$ is true at the current
world iff $\varphi$ is \emph{possible}, i.e. true in at least some world
accessible from the current world.

>>>>

\begin{itemize}
  \item We define a \emph{deep embedding} of modal logic syntax in Coq and
    define its Kripke semantics.
  \item We model the Hilbert deductive system as a Coq inductive type that is
    parameterized by different axiom systems like K, D, B, T, K4, K5, S4, S5 and GL,
    and prove its soundness with respect to Kripke semantics.
    In such, we also formalize several frame correspondence results.
  \item As an example of usage, we show a proof of L\"ob's theorem using the deductive system
    developed in our library.
\end{itemize}

@@

Coq is a proof assistant based on the calculus of inductive
constructions (CIC)~\cite{manual_cic}, a higher-order typed
$\lambda$-calculus extended with inductive definitions.  Theorem
proving in Coq follows the ideas of the so-called
``BHK-in\-ter\-pre\-ta\-tion''\footnote{Abbreviation of Brouwer, Heyting,
  Kolmogorov, de Bruijn and Martin-L\"of Correspondence. This is also
  known as the Curry-Howard isomorphism.}, where types represent
logical formulas, $\lambda$-terms represent proofs
\cite{Bertot10} and the task of checking if a piece of text is a
proof of a given formula corresponds to checking if the term that
represents the proof has the type corresponding to the given formula.

However, writing a proof term whose type is that of a logical formula
can be challenging, even for very simple propositions.  In order to
make the writing of complex proofs easier, Coq provides
\emph{tactics}, which are commands that can be used to construct proof
terms in a more user-friendly way.

@@

\paragraph{An overview of modal logic.} Modal logics are designed for reasoning
about truth across various --- abstract --- worlds. In such logics, a
proposition may be true in some world, but false in another one. The systems of
modal logic considered in this work extend traditional propositional logic
with two operators on propositions: $\Box$ (box) and $\Diamond$ (diamond).
The syntax for modal logic formulas are defined by the following context free
grammar (where $\circ \in\{\land,\lor,\to,\leftrightarrow\}$ and
$\mathcal{P} = \{A,B,C,...\}$ is the enumerable set of propositional variables):
\[
  \begin{array}{lcl}
    \varphi & ::= & \mathcal{P}\,|\,\neg\varphi\,|\,\varphi\circ\varphi\,|\,\Box\varphi\,|\,\Diamond\varphi
  \end{array}  
\]
Intuitively, the semantics of the operator $\Box$ is similar to an universal
quantifier, since it expresses \emph{necessity}: $\Box\,\varphi$ is true in the
current world iff $\varphi$ is true in every world accessible from the current
one. Similarly, the semantics of operator
$\Diamond$ is that a formula $\Diamond\varphi$ is true in the current
world iff $\varphi$ is \emph{possible}, i.e. true in at least some world
accessible from the current world.
